Studies of quantum dots

### 2016-11-11 Manuscript/paper.tex
Fix some references

This article is organized as follows: Section \ref{sec:formalism} introduces
first (\ref{subsec:modelHamiltonian}) the Hamiltonian we use to model circular
quantum dots, and gives afterwards an overview of the IM-SRG
(\ref{subsec:SRG}) and DMC (\ref{subsec:DMC}) formalisms.  Our results are
presented in Section \ref{sec:results}.  We point out the differences arising
from the use of two different generators (Wegner and White generator), and
give an overview of the ground state energies up to 42 particles. In
particular, we compare the results obtained with five different \textit{ab
  initio} methods: IM-SRG(2), DMC, Hartree-Fock, FCI and CC. Section
\ref{sec:conclusions} concludes our work and gives perspectives for future
work.

@@

\subsubsection{Wegner's canonical generator}

@@

\subsubsection{White's generator}

@@

Making use of the assumptions in Eq.\ \eqref{eq:omega-condition1} and \eqref{eq:omega-condition2}, one can derive from the Schr\"odinger equation the generalized Bloch equation, the principal equation of QDPT:
\begin{gather*}
  [\hat \Omega, \hat H^{\text o}] =
  (1 - \hat \Omega) \hat V \Omega
\end{gather*}
where $\hat V \equiv \hat H - \hat H^{\text o}$ is the perturbation.  The
commutator on the left may be ``inverted'' using the resolvent approach
(\cite{shavitt2009many}, p.\ 50), resulting in:
\begin{align*}
  \hat Q \Omega \hat P_u =
  \hat R_u (1 - \hat \Omega) \hat V \Omega \hat P_u
\end{align*}
where $\hat R_u \equiv \hat Q (E_u - \hat Q \hat H^{\text o} \hat Q)^{-1} \hat Q$ is the resolvent and $\hat P_u$ is the projection operator that projects any state onto $\Psi^{\text o}_u$.  As is typical in perturbation theory, we assume $\hat \Omega$ can be expanded as a series of the increasing order as measured by the ``power'' of the perturbation $\hat V$:
\begin{align*}
  \hat \Omega = \hat P +
  \hat Q\bigl(\hat \Omega_1 + \hat \Omega_2 + \cdots\bigr) \hat P
\end{align*}
This leads to a recursion relation of $\hat \Omega$ that enables $\hat \Omega$ to be calculated up to any order, as least in principle.  Up to third order, we have: [[double-check]]
\begin{align*}
  &\hat \Omega_1 \hat P_u = \hat R_u \hat V \hat P_u \\
  &\hat \Omega_2 \hat P_u =
    \hat R_u \biggl(
    \hat V \hat R_u
    - \sum_v \hat R_v \hat V \hat P_v
    \biggr) \hat V \hat P_u \\
  &\hat \Omega_3 \hat P_u =
    \hat R_u \biggl(
    \hat V \hat R_u \hat V \hat R_u
    - \hat V \hat R_u \sum_v \hat R_v \hat V \hat P_v \\
  &\qquad\qquad
    - \sum_v \hat R_v \hat V \hat P_v \hat V \hat R_u
    - \sum_v \hat R_v \hat V \hat R_v \hat V \hat P_v \\
  &\qquad\qquad
    + \sum_v \hat R_v \sum_w \hat R_w \hat V \hat P_w \hat V \hat P_v
    \biggr) \hat V \hat P_u
\end{align*}
Until now, the solution is described purely in terms of abstract operators in Hilbert space.  To make this concrete and relevant for many-body physics, we must make some additional assumptions.  First, we make use of the single-particle basis as we have for HF and IM-SRG.  Secondly, we assume the set of reference states $\Psi^{\text o}_u$ are Slater determinants of the single-particle states.

@@

\subsection{Diffusion Monte Carlo}

@@

The Green's function introduced in Eq.~(\ref{eq:DMC_GreenFuncRevealed}) has singularities due to the the Coulomb interaction term in $\hat{H}$. Due to Pauli exclusion these are not physical, nevertheless, in numerical implementations they are unpredictable and undesirable, hence we want to avoid them.

@@

The Green's function from Eq.~(\ref{DMC_IS_GF_raw}) has been split into two for practical reasons; one part containing the kinetic terms, more specifically this represents a Fokker-Planck diffusion process, and a second part which is commonly referred to as a \textit{branching function}, which is an effective reweighing of a positions contribution to the overall density. Excluding branching we will simply sample from the density of the trial state.

In order dampen the errors introduced by splitting the Green's function, the time step $\delta\tau$ should be small. A value in the range of $10^{-3}$ to $10^{-4}$ is used in this work. High variance in energy samples requires a smaller time step to keep the branching controlled, since the fluctuations of the two are connected by Eq.~(\ref{eq:DMC_IS_GF}).

The name branching arise from the fact that in DMC, the weights are implemented as a probability of a given walker to copy or delete itself $n$ times, where $\langle n \rangle$ equals the branching term of Eq.~(\ref{eq:DMC_IS_GF}). Let $G_b$ denote the value of the branching function at a given position. Correct weighting is then ensured by guaranteeing $m = \max\left(\mathrm{floor}\left(G_b\right)-1, 0\right)$ branches with a probability $G_b - m$ of getting an additional branch. A branch is an identical copy of the corresponding walker, with an identical past but with a different future. When the number of requested branches is zero, the walker is removed from the simulation.

Initially, $N_w$ walkers are initialized to span $|\Psi_T(\mathbf{r})|^2$. This is ensured by using the Metropolis Algorithm. Moreover, importance sampling is ensured by sampling according to the kinetic term of  Eq.~(\ref{eq:DMC_IS_GF_raw}) \textit{cite solution langevin}

@@

The Slater determinant contains eigenstates $\phi_n(\mathbf{r})$ of Eq.~(\ref{eq:potential})
with a variational parameter $\alpha$ scaling the oscillator frequency $\omega$
such that $\omega\to\alpha\omega$ in all solutions:

@@

\begin{table}
\begin{center}
\caption{Ground state energy (in atomic units) of a circular quantum
  dot, with $N$ particles and oscillator frequency $\omega$. The
  IM-SRG(2) results are given for $R=20$ shells (except for the $N=56$
  results which are for $R=18$), employing a Hartree-Fock basis, a
  bare Coulomb interaction and White's generator. For $N=42$ and
  $\omega=0.1$ results are presented up to $R=22$ major shells. The
  results labelled IM-SRG(2)$_{R\rightarrow \infty}$ are obtained
  using the extrapolation formula of Eq.().}
\label{tab:SRGDMC-R20}
\begin{tabular}{ccccc}
\hline\hline
$N$ & $\omega$ & IM-SRG(2) &IM-SRG(2)$_{R\rightarrow \infty}$ & DMC \\
\hline
6 & 1.0 & 20.1583 & 20.1468   & 20.1593(1) \\
& 0.5&11.7690  & 11.7631 &11.7848(1) \\
& 0.28& 7.5713 &   7.5695 &7.6002(1) \\
& 0.1& 3.4962  &  3.4977 &3.5539(1)\\
\hline
12&1.0 & 65.7280 & 65.6990  & 65.7001(1) \\
&0.5 &39.1631 &  39.1473 & 39.1596(1)\\
&  0.28&25.6212 &  25.6140& 25.6358(1)\\
&  0.1&12.2223 &  12.2234 & 12.2698(1)\\
\hline
20& 1.0&155.9737 & 155.9130  & 155.8822(1)\\
&  0.5&93.9219 &   93.8853 &93.8752(1)\\
&  0.28&61.94362 & 61.9240 &61.9268(1) \\
&  0.1&29.95263 &29.9565  & 29.9779(1)\\
\hline
30& 1.0& 308.7673 & 308.6053 & 308.5627(2) \\
 & 0.5& 187.1671& 186.8340 &187.0426(2)\\
 & 0.28&124.0410 & 123.7171 & 123.9683(2) \\
 & 0.1& 60.4319& 60.1158 & 60.4205(2)\\
\hline
42& 1.0&543.3398 & 543.3304 & 542.9428(8) \\
 & 0.5&330.8885 & 330.8425 & 330.6306(2)\\
 & 0.28&220.0226 & 219.9534y &219.8426(2) \\
 & 0.1& 107.7851&  106.8074& 107.6389(2)\\
 \hline
56& 1.0& 880.4159 &  & 879.3986(6) \\
 & 0.5& 538.8199 & & 537.353(2)\\
 & 0.28&360.83689  &  & 358.145(2) \\
\hline
\end{tabular}
\end{center}
\end{table}

In Figure \ref{fig:gs} we show the calculated ground state energies from
Hartree-Fock, second-order (M\o ller-Plesset) perturbation theory (MPPT), and
IM-SRG.  The exact result from diffusion Monte Carlo (DMC) is also shown as a
dashed line in systems where it is known \cite{PhysRevB.84.115302}.

@@

For our addition and removal energy calculations, the results are summarized
in Figure \ref{fig:add} and Figure \ref{fig:rm} respectively.  The figures
show the the addition/removal energies for Hartree-Fock, combined with IM-SRG
and/or QDPT (Hartree-Fock is always included, as usual) [[better way to
explain this? the figure labels are somewhat misleading too]].  The DMC
results are also shown via dashed lines where they are known.

@@

@@

\begin{acknowledgments}
 This work was supported by the
National Science Foundation Grant No.~PHY-1404159
(Michigan State University) and by the Research
Council of Norway under contract ISP-Fysikk/216699.
\end{acknowledgments}

@@

\documentclass[aps,twocolumn,showpacs,floatfix,nofootinbib,preprintnumbers,superscriptaddress,amsmath,amssymb]{revtex4-1}

\usepackage{graphicx}
\usepackage{epsfig}
\usepackage{bm}
\usepackage{color}
\usepackage{float}
\usepackage{dcolumn}
\usepackage{subfigure}
\usepackage{multirow}

\section{Results}
\label{sec:results}
In this section, we discuss the IM-SRG results for the ground-state
energies of closed-shell systems. In subsection \ref{subsec:Wegner},
we start with Wegner's canonical generator. Numerical instabilities in
the integration process, as well as a rather slow convergence as
function of the number of oscillator shells, motivate the use of a
Hartree-Fock basis and the Coulomb interaction. Since higher
correlations still lead to stiff equation systems, we change in
subsection \ref{subsec:White} to White's generator. In particular, we
give an overview of the IM-SRG(2) ground state energies for
closed-shell systems up to $N=42$ particles and compare with other
\textit{ab initio} many-body methods.\\ In all our calculations, we
perform the integration of the flow equations using the algorithm
developed by Shampine and Gordon \cite{shampine1975computer}. This
algorithm solves linear systems of differential equations based on the
implicit Adams method, which is a multi-step method of variable
order. Source code is freely available \cite{odesolver}.

\section{Conclusions}
\label{sec:conclusions}
We have used IM-SRG to study the ground state energy of circular,
two-dimensional quantum dots up to $N=42$ particles. We utilized two
different generators, Wegner's and White's one, and realized that
Wegner's generator results in stiff equation systems, whereas White's
generator has much better numerics and is computationally more
efficient. Moreover, we found out that the use of an Hartree-Fock
basis improves numerical stability enormously, whereas analogous to
previous FCI \cite{Kvaalcode} and CCSD \cite{PhysRevB.84.115302}
calculations, the use of an effective interaction improves convergence
of the ground state energy as function of the number of oscillator
shells.  Our IM-SRG(2) results are in excellent agreement with other
many-body methods, in particular DMC, which can be considered as more
or less exact benchmark. For more than six particles, we lie in all
cases closer to the DMC result than corresponding CCSD calculations
do.

%
\begin{acknowledgments}
  We thank Christoffer Hirth, Simen Kvaal and Veronica Olsen for
  several discussions. This work was supported by the Research Council
  of Norway under contract ISP-Fysikk/216699. This research used
  computational resources of the Notur project in Norway.
\end{acknowledgments}

\bibliography{srg}
\bibliographystyle{apsrev4-1}

\end{document}

>>>>

This article is organized as follows: Section \ref{sec:formalism} introduces
first (\ref{subsec:modelHamiltonian}) the Hamiltonian we use to model circular
quantum dots, and gives afterwards an overview of the IM-SRG
(\ref{subsec:imsrgmethod}) and DMC (\ref{subsec:DMC}) formalisms.  Our results are
presented in Section \ref{sec:results}.  We point out the differences arising
from the use of two different generators (Wegner and White generator), and
give an overview of the ground state energies up to 42 particles. In
particular, we compare the results obtained with five different \textit{ab
  initio} methods: IM-SRG(2), DMC, Hartree-Fock, FCI and CC. Section
\ref{sec:conclusions} concludes our work and gives perspectives for future
work.

@@

\subsubsection{Wegner's canonical generator}
\label{subsec:Wegner}

@@

\subsubsection{White's generator}
\label{subsec:White}

@@

Making use of the assumptions in  \eqref{eq:omega-condition1} and \eqref{eq:omega-condition2}, one can derive from the Schr\"odinger equation the generalized Bloch equation, the principal equation of QDPT:
\begin{gather*}
  [\hat \Omega, \hat H^{\text o}] =
  (1 - \hat \Omega) \hat V \Omega
\end{gather*}
where $\hat V \equiv \hat H - \hat H^{\text o}$ is the perturbation.  The
commutator on the left may be ``inverted'' using the resolvent approach
(\cite{shavitt2009many}, p.\ 50), resulting in:
\begin{align*}
  \hat Q \Omega \hat P_u =
  \hat R_u (1 - \hat \Omega) \hat V \Omega \hat P_u
\end{align*}
where $\hat R_u \equiv \hat Q (E_u - \hat Q \hat H^{\text o} \hat Q)^{-1} \hat Q$ is the resolvent and $\hat P_u$ is the projection operator that projects any state onto $\Psi^{\text o}_u$.  As is typical in perturbation theory, we assume $\hat \Omega$ can be expanded as a series of the increasing order as measured by the ``power'' of the perturbation $\hat V$:
\begin{align*}
  \hat \Omega = \hat P +
  \hat Q\bigl(\hat \Omega_1 + \hat \Omega_2 + \cdots\bigr) \hat P
\end{align*}
This leads to a recursion relation of $\hat \Omega$ that enables $\hat \Omega$ to be calculated up to any order, as least in principle.  Up to third order, we have: [[double-check]]
\begin{align*}
  &\hat \Omega_1 \hat P_u = \hat R_u \hat V \hat P_u \\
  &\hat \Omega_2 \hat P_u =
    \hat R_u \biggl(
    \hat V \hat R_u
    - \sum_v \hat R_v \hat V \hat P_v
    \biggr) \hat V \hat P_u \\
  &\hat \Omega_3 \hat P_u =
    \hat R_u \biggl(
    \hat V \hat R_u \hat V \hat R_u
    - \hat V \hat R_u \sum_v \hat R_v \hat V \hat P_v \\
  &\qquad\qquad
    - \sum_v \hat R_v \hat V \hat P_v \hat V \hat R_u
    - \sum_v \hat R_v \hat V \hat R_v \hat V \hat P_v \\
  &\qquad\qquad
    + \sum_v \hat R_v \sum_w \hat R_w \hat V \hat P_w \hat V \hat P_v
    \biggr) \hat V \hat P_u
\end{align*}
Until now, the solution is described purely in terms of abstract operators in Hilbert space.  To make this concrete and relevant for many-body physics, we must make some additional assumptions.  First, we make use of the single-particle basis as we have for HF and IM-SRG.  Secondly, we assume the set of reference states $\Psi^{\text o}_u$ are Slater determinants of the single-particle states.

@@

\subsection{Diffusion Monte Carlo}
\label{subsec:DMC}

@@

The Green's function introduced in \eqref{eq:DMC_GreenFuncRevealed} has singularities due to the the Coulomb interaction term in $\hat{H}$. Due to Pauli exclusion these are not physical, nevertheless, in numerical implementations they are unpredictable and undesirable, hence we want to avoid them.

@@

The Green's function from \eqref{eq:DMC_IS_GF_raw} has been split into two for practical reasons; one part containing the kinetic terms, more specifically this represents a Fokker-Planck diffusion process, and a second part which is commonly referred to as a \textit{branching function}, which is an effective reweighing of a positions contribution to the overall density. Excluding branching we will simply sample from the density of the trial state.

In order dampen the errors introduced by splitting the Green's function, the time step $\delta\tau$ should be small. A value in the range of $10^{-3}$ to $10^{-4}$ is used in this work. High variance in energy samples requires a smaller time step to keep the branching controlled, since the fluctuations of the two are connected by \eqref{eq:DMC_IS_GF}.

The name branching arise from the fact that in DMC, the weights are implemented as a probability of a given walker to copy or delete itself $n$ times, where $\langle n \rangle$ equals the branching term of \eqref{eq:DMC_IS_GF}. Let $G_b$ denote the value of the branching function at a given position. Correct weighting is then ensured by guaranteeing $m = \max\left(\mathrm{floor}\left(G_b\right)-1, 0\right)$ branches with a probability $G_b - m$ of getting an additional branch. A branch is an identical copy of the corresponding walker, with an identical past but with a different future. When the number of requested branches is zero, the walker is removed from the simulation.

Initially, $N_w$ walkers are initialized to span $|\Psi_T(\mathbf{r})|^2$. This is ensured by using the Metropolis Algorithm. Moreover, importance sampling is ensured by sampling according to the kinetic term of \eqref{eq:DMC_IS_GF_raw} [[cite solution langevin]]

@@

The Slater determinant contains eigenstates $\phi_n(\mathbf{r})$ of the 2-dimensional harmonic oscillator
with a variational parameter $\alpha$ scaling the oscillator frequency $\omega$
such that $\omega\to\alpha\omega$ in all solutions:

@@

\begin{table}
\begin{center}
\caption{Ground state energy (in atomic units) of a circular quantum
  dot, with $N$ particles and oscillator frequency $\omega$. The
  IM-SRG(2) results are given for $R=20$ shells (except for the $N=56$
  results which are for $R=18$), employing a Hartree-Fock basis, a
  bare Coulomb interaction and White's generator. For $N=42$ and
  $\omega=0.1$ results are presented up to $R=22$ major shells. The
  results labelled IM-SRG(2)$_{R\rightarrow \infty}$ are obtained
  using the extrapolation formula of [[equation?]].}
\label{tab:SRGDMC-R20}
\begin{tabular}{ccccc}
\hline\hline
$N$ & $\omega$ & IM-SRG(2) &IM-SRG(2)$_{R\rightarrow \infty}$ & DMC \\
\hline
6 & 1.0 & 20.1583 & 20.1468   & 20.1593(1) \\
& 0.5&11.7690  & 11.7631 &11.7848(1) \\
& 0.28& 7.5713 &   7.5695 &7.6002(1) \\
& 0.1& 3.4962  &  3.4977 &3.5539(1)\\
\hline
12&1.0 & 65.7280 & 65.6990  & 65.7001(1) \\
&0.5 &39.1631 &  39.1473 & 39.1596(1)\\
&  0.28&25.6212 &  25.6140& 25.6358(1)\\
&  0.1&12.2223 &  12.2234 & 12.2698(1)\\
\hline
20& 1.0&155.9737 & 155.9130  & 155.8822(1)\\
&  0.5&93.9219 &   93.8853 &93.8752(1)\\
&  0.28&61.94362 & 61.9240 &61.9268(1) \\
&  0.1&29.95263 &29.9565  & 29.9779(1)\\
\hline
30& 1.0& 308.7673 & 308.6053 & 308.5627(2) \\
 & 0.5& 187.1671& 186.8340 &187.0426(2)\\
 & 0.28&124.0410 & 123.7171 & 123.9683(2) \\
 & 0.1& 60.4319& 60.1158 & 60.4205(2)\\
\hline
42& 1.0&543.3398 & 543.3304 & 542.9428(8) \\
 & 0.5&330.8885 & 330.8425 & 330.6306(2)\\
 & 0.28&220.0226 & 219.9534y &219.8426(2) \\
 & 0.1& 107.7851&  106.8074& 107.6389(2)\\
 \hline
56& 1.0& 880.4159 &  & 879.3986(6) \\
 & 0.5& 538.8199 & & 537.353(2)\\
 & 0.28&360.83689  &  & 358.145(2) \\
\hline
\end{tabular}
\end{center}
\end{table}

In Figure [[fig:gs]] we show the calculated ground state energies from
Hartree-Fock, second-order (M\o ller-Plesset) perturbation theory (MPPT), and
IM-SRG.  The exact result from diffusion Monte Carlo (DMC) is also shown as a
dashed line in systems where it is known \cite{PhysRevB.84.115302}.

@@

For our addition and removal energy calculations, the results are summarized
in Figure [[fig:add]] and Figure [[fig:rm]] respectively.  The figures
show the the addition/removal energies for Hartree-Fock, combined with IM-SRG
and/or QDPT (Hartree-Fock is always included, as usual) [[better way to
explain this? the figure labels are somewhat misleading too]].  The DMC
results are also shown via dashed lines where they are known.

@@

[[begin unmerged]]

In this section, we discuss the IM-SRG results for the ground-state
energies of closed-shell systems. In subsection \ref{subsec:Wegner},
we start with Wegner's canonical generator. Numerical instabilities in
the integration process, as well as a rather slow convergence as
function of the number of oscillator shells, motivate the use of a
Hartree-Fock basis and the Coulomb interaction. Since higher
correlations still lead to stiff equation systems, we change in
subsection \ref{subsec:White} to White's generator. In particular, we
give an overview of the IM-SRG(2) ground state energies for
closed-shell systems up to $N=42$ particles and compare with other
\textit{ab initio} many-body methods.\\ In all our calculations, we
perform the integration of the flow equations using the algorithm
developed by Shampine and Gordon \cite{shampine1975computer}. This
algorithm solves linear systems of differential equations based on the
implicit Adams method, which is a multi-step method of variable
order. Source code is freely available \cite{odesolver}.

[[end unmerged]]

@@

[[begin unmerged]]

We have used IM-SRG to study the ground state energy of circular,
two-dimensional quantum dots up to $N=42$ particles. We utilized two
different generators, Wegner's and White's one, and realized that
Wegner's generator results in stiff equation systems, whereas White's
generator has much better numerics and is computationally more
efficient. Moreover, we found out that the use of an Hartree-Fock
basis improves numerical stability enormously, whereas analogous to
previous FCI \cite{Kvaalcode} and CCSD \cite{PhysRevB.84.115302}
calculations, the use of an effective interaction improves convergence
of the ground state energy as function of the number of oscillator
shells.  Our IM-SRG(2) results are in excellent agreement with other
many-body methods, in particular DMC, which can be considered as more
or less exact benchmark. For more than six particles, we lie in all
cases closer to the DMC result than corresponding CCSD calculations
do.

[[end unmerged]]

\begin{acknowledgments}
  This work was supported by the
  National Science Foundation Grant No.~PHY-1404159
  (Michigan State University) and by the Research
  Council of Norway under contract ISP-Fysikk/216699.

  [[begin unmerged]]

  We thank Christoffer Hirth, Simen Kvaal and Veronica Olsen for
  several discussions. This work was supported by the Research Council
  of Norway under contract ISP-Fysikk/216699. This research used
  computational resources of the Notur project in Norway.

  [[end unmerged]]
\end{acknowledgments}

@@
